---
title: "R Explorating"
author: "Rony Silva"
date: "25/10/2019"
output: github_document
---

Hi, I am Ronyberque C. Silva, a student of Msc. on Informatics on University Federal of Alagoas - UFAL, on Brazil. In this article, I show the use of some functions of R.

### Choose Dataset
The Dataset choose was a dataset that contains all sweepstakes of Mega-Sena (a lotery of Brazil).
The dataset was downloaded from [Kaggle.com](https://www.kaggle.com) and has tunned to this example.
Link to download. [https://www.kaggle.com/viniciusbbizarri/sorteiosmegasena](https://www.kaggle.com/viniciusbbizarri/sorteiosmegasena).

```{r}
library(dplyr)
library(ggplot2)
```

### Define path of Datasets
Below i define the path of used datasets
```{r}
path <- "/home/rony/Documentos/data-analysis/Datasets/"
```

### Import Dataset
R allows that the developer can import data from a CSV and other extensions. CSV files without header can take ignore header using the param 'header=FALSE'. On default, 'header=TRUE', like you see below.

```{r}
sorteios = read.csv(paste(path,"sorteios.csv",sep=""),header=TRUE)
```

I'm specialy interested on obtain informations related to winners of Sena. Bellow, i am start to refine the content of dataset choised.

## Explorating dataset through R functions

### str Function
The use of 'str' function, allows that the structure of an object R can be exposed. This function allows show basically the structure of the assigned object but in resume.

```{r}
str(sorteios)
```

### glimpse Function
The 'glimpse' function is like an 'str', but, basically with the difference that 'glimpse' try to show all structure of assign object.

- On the show, rows are attributes and columns are data of these attributes.
- Glimpse allows that the developer can inform of the second param of function, the limit of data to show. On default, the 'width' param (second param), is the width of the console. 

```{r}
glimpse(sorteios)
```


Then, how can be visualized, the chosen dataset contains 11 features (columns) and 2278 points of data (rows). 2278 registers of sweepstakes since 1996 to final of 2018.


### table Function

The function 'table' allows that the developer can show features of dataset using a filter, for instance. 
One of the features of the dataset 'sorteios', is the 'Ganhadores_Sena'. How can see bellow.
R allows summarizing information of a specific feature only using '$' and passing the feature name (in this case, Ganhadores_Sena) existent on dataset.

```{r}
table(sorteios$Ganhadores_Sena)
```

How can see, many observations contains instances that not have winners, and how i am interested only sweepstakes that contains leastwise one winner on Sena, then i decided filter my object sorteios and return only instances that contains Ganhadores_Sena > 0.

```{r}
onlyWithSenaWinners <- filter(sorteios, Ganhadores_Sena > 0)
```

The result can be viewed bellow...

```{r}
table(onlyWithSenaWinners$Ganhadores_Sena)
```

How can be viewed above, 357 observations had one winner on "Sena", for instance. The representation above stills need detail the total of winners, because that filter applied only show the total of instances that are framed on the filter. Below a sum of data points to each instance is applied.

### Ploting Graphs
On R, the graph plotting is easy to be used. On r-graph-gallery ([https://www.r-graph-gallery.com/]), a lot of graphs and how can be used, can be viewed.

Below, I use a graph bar to confront the total of winners on "Sena" versus "Quina".

```{r}
totalWinnersSena = sum(onlyWithSenaWinners$Ganhadores_Sena)
totalWinnersQuina = sum(onlyWithSenaWinners$Ganhadores_Quina)
totalWinnersQuadra = sum(onlyWithSenaWinners$Ganhadores_Quadra)


barplot(height=c(totalWinnersSena,totalWinnersQuina), names=c("Sena","Quina"), main="Winners of Sena versus Quina")

```
### Correlations between Sena, Quina and Quadra
The correlation test is used to verify the association between two or more variables. Here is used to verify the correlation between winners of Sena and Quina.

```{r}
  correlation <- cor.test(onlyWithSenaWinners$Ganhadores_Sena, onlyWithSenaWinners$Ganhadores_Quina,  method="kendall")
  correlation
```

Then, how can be viewed above, the correlation between winners of Sena and Quina is very low (More winners of quina does not necessarily increase the number of seine winners on sena).

```{r}
  correlation2 <- cor.test(onlyWithSenaWinners$Ganhadores_Quina, onlyWithSenaWinners$Ganhadores_Quadra,  method="kendall")
  correlation2
```

On correlation between Quina and Quadra the result is more animator. 

So that's it person, until next time






