---
title: "R Explorating"
author: "Rony Silva"
date: "25/10/2019"
output: github_document
---

### Choose Dataset
The Dataset choose was a dataset that contains all sweepstakes of Mega-Sena (a lotery of Brazil).
The dataset was downloaded from [Kaggle.com](https://www.kaggle.com).
Link to download. [https://www.kaggle.com/viniciusbbizarri/sorteiosmegasena](https://www.kaggle.com/viniciusbbizarri/sorteiosmegasena).

```{r}
library(dplyr)
```

### Define path of Datasets
Below i define the path of used datasets
```{r}
path <- "/home/rony/Documentos/data-analysis/Datasets/"
```

### Import Dataset
R allows that the developer can import data from a CSV and other extensions. CSV files without header can take ignore header using the param 'header=FALSE'. On default, 'header=TRUE', like you see below.

```{r}
sorteios = read.csv(paste(path,"sorteios.csv",sep=""),header=TRUE)
```

I'm specialy interested on obtain informations related to winners on one state of Brazil. Bellow, i am start to refine the content of dataset choised.

## Explorating dataset through R functions

### str Function
The use of 'str' function, allows that the structure of an object R can be exposed. This function allows show basically the structure of the assigned object but in resume.

```{r}
str(sorteios)
```

### glimpse Function
The 'glimpse' function is like an 'str', but, basically with the difference that 'glimpse' try to show all structure of assign object.

- On the show, rows are attributes and columns are data of these attributes.
- Glimpse allows that the developer can inform of the second param of function, the limit of data to show. On default, the 'width' param (second param), is the width of the console. 

```{r}
glimpse(sorteios)
```


Then, how can be visualized, the chosen dataset contains 23 features (columns) and 2278 points of data (rows). 2278 registers of sweepstakes since 1996 to final of 2018.


### table Function

The function 'table' allows that the developer can show features of dataset using a filter, for instance. 
One of the features of the dataset 'sorteios', is the 'UF' of winner. How can see bellow.
R allows summarizing information of a specific feature only using '$' and passing the feature name (in this case, UF) existent on dataset.

```{r}
table(sorteios$UF)
```

How can see, SP contains 202 winner, but how i am interested only sweepstakes that contains a
'#' state, then i decided filter my object sorteios and return only data poins that contains 
'#' a UF (1600 not contains).'

```{r}
onlySorteiosWithUF <- filter(sorteios, UF != "")
```

The result can be viewed bellow... Observe the first column of the table.
This represent the total data points that contains winners of one state.

```{r}
table(onlySorteiosWithUF$UF)
```

